---
title: "Mid-Term Draft 02"
author: "Konrad Bailey"
date: "2024-03-20"
output: pdf_document
eval: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load required libraries
```{r}
library(readxl)
library(httr)
library(jsonlite)
library(tidyverse)
library(dplyr)
library(tidyr)
```

```{r}
USDA_UI.data <- read.csv("~/Desktop/USDAtoUI.csv")
```

```{r}
head(USDA_UI.data)
```


```{r}
DOEawards.data <- read.csv("~/Desktop/DOEawards.csv")
```

### Filter city by Moscow in DOE data
```{r}
DOE_UI.data <- DOEawards.data %>%
  filter(City == 'Moscow')
```

```{r}
head(DOE_UI.data)
```

### Fetch NSF data with API
```{r} 
#| eval: false


library(readxl)


# Base URL for the API
base_url <- "https://www.research.gov/awardapi-service/v1/awards.json?awardeeName=%22regents+of+the+university+of+idaho%22"

printFields <- "rpp,offset,id,agency,awardeeCity,awardeeCountryCode,awardeeDistrictCode,awardeeName,awardeeStateCode,awardeeZipCode,cfdaNumber,coPDPI,date,startDate,expDate,estimatedTotalAmt,fundsObligatedAmt,ueiNumber,fundProgramName,parentUeiNumber,pdPIName,perfCity,perfCountryCode,perfDistrictCode,perfLocation,perfStateCode,perfZipCode,poName,primaryProgram,transType,title,awardee,poPhone,poEmail,awardeeAddress,perfAddress,publicationResearch,publicationConference,fundAgencyCode,awardAgencyCode,projectOutComesReport,abstractText,piFirstName,piMiddeInitial,piLastName,piEmail"

# Initialize an empty data frame to store results
NSFtoUI <- tibble()

# Number of results per page (as per API settings)
results_per_page <- 25

# Variable to keep track of the current page number
current_page <- 1

# Variable to control the loop
keep_going <- TRUE

while(keep_going) {
    # Calculate the offset for the current page
    offset <- (current_page - 1) * results_per_page + 1

    # Construct the full URL with offset
    url <- paste0(base_url, "&offset=", offset, "&printFields=", printFields)

    # Make the API call
    response <- GET(url)

    # Check if the call was successful
    if (status_code(response) == 200) {
        # Extract and parse the JSON data
        json_data <- content(response, type = "text", encoding = "UTF-8")
        parsed_data <- fromJSON(json_data, flatten = TRUE)

        # Extract the 'award' data and add to the all_awards data frame
        awards_data <- parsed_data$response$award
        NSFtoUI <- bind_rows(NSFtoUI, as_tibble(awards_data))

        # Debug: Print the current page number and number of awards fetched
        print(paste("Page:", current_page, "- Awards fetched:", length(awards_data$id)))

        # Check if the current page has less than results_per_page awards, then it's the last page
        if (length(awards_data$id) < results_per_page) {
            keep_going <- FALSE
        } else {
            current_page <- current_page + 1
        }
    } else {
        print(paste("Failed to fetch data: Status code", status_code(response)))
        keep_going <- FALSE
    }
}
```

```{r}
head(projects_df)
```


# Creating My New Dataset:

```{r}
# Prepare DOE_UI.data
DOE_UI_prepared <- DOE_UI.data %>%
  select(PI_Name = PI, 
         Award_Amount = Amount.Awarded.to.Date, 
         Start_Date = Start.Date, 
         End_Date = End.Date) %>%
  mutate(Award_Amount = as.numeric(gsub("[^0-9.-]", "", Award_Amount)),
         Start_Date = as.Date(Start_Date, format = "%Y-%m-%d"),
         End_Date = as.Date(End_Date, format = "%Y-%m-%d"))

# Prepare NSFtoUI
NSFtoUI_prepared <- NSFtoUI %>%
  select(PI_Name = pdPIName, 
         Award_Amount = estimatedTotalAmt, 
         Start_Date = startDate, 
         End_Date = expDate) %>%
  mutate(Award_Amount = as.numeric(gsub("[^0-9.-]", "", Award_Amount)),
         Start_Date = as.Date(Start_Date, format = "%Y-%m-%d"),
         End_Date = as.Date(End_Date, format = "%Y-%m-%d"))

# Prepare projects_df
projects_df_prepared <- projects_df %>%
  select(PI_Name = contact_pi_name, 
         Award_Amount = award_amount, 
         Start_Date = project_start_date, 
         End_Date = project_end_date) %>%
  mutate(Award_Amount = as.numeric(gsub("[^0-9.-]", "", Award_Amount)),
         Start_Date = as.Date(Start_Date, format = "%Y-%m-%d"),
         End_Date = as.Date(End_Date, format = "%Y-%m-%d"))

# Combine the prepared datasets
UI_Grant_Data <- bind_rows(DOE_UI_prepared, NSFtoUI_prepared, projects_df_prepared)
```

```{r}
head(UI_Grant_Data)
```

# Test:


# Prepare DOE_UI.data with Funding_Agency
DOE_UI_prepared <- DOE_UI.data %>%
  select(PI_Name = PI, 
         Award_Amount = Amount.Awarded.to.Date, 
         Start_Date = Start.Date, 
         End_Date = End.Date) %>%
  mutate(Award_Amount = as.numeric(gsub("[^0-9.-]", "", Award_Amount)),
         Start_Date = as.Date(Start_Date, format = "%Y-%m-%d"),
         End_Date = as.Date(End_Date, format = "%Y-%m-%d"),
         Funding_Agency = "DOE")

# Prepare NSFtoUI with Funding_Agency
NSFtoUI_prepared <- NSFtoUI %>%
  select(PI_Name = pdPIName, 
         Award_Amount = estimatedTotalAmt, 
         Start_Date = startDate, 
         End_Date = expDate) %>%
  mutate(Award_Amount = as.numeric(gsub("[^0-9.-]", "", Award_Amount)),
         Start_Date = as.Date(Start_Date, format = "%Y-%m-%d"),
         End_Date = as.Date(End_Date, format = "%Y-%m-%d"),
         Funding_Agency = "NSF")

# Prepare projects_df with Funding_Agency
projects_df_prepared <- projects_df %>%
  select(PI_Name = contact_pi_name, 
         Award_Amount = award_amount, 
         Start_Date = project_start_date, 
         End_Date = project_end_date) %>%
  mutate(Award_Amount = as.numeric(gsub("[^0-9.-]", "", Award_Amount)),
         Start_Date = as.Date(Start_Date, format = "%Y-%m-%d"),
         End_Date = as.Date(End_Date, format = "%Y-%m-%d"),
         Funding_Agency = "NIH")

# Combine the prepared datasets
UI_Grant_Data <- bind_rows(DOE_UI_prepared, NSFtoUI_prepared, projects_df_prepared)



```{r}
#head(UI_Grant_Data)
```

```{r}
# Prepare NSFtoUI with corrected date handling and Funding_Agency
NSFtoUI_prepared <- NSFtoUI %>%
  select(
    PI_Name = pdPIName, 
    Award_Amount = estimatedTotalAmt, 
    Start_Date = startDate, 
    End_Date = expDate
  ) %>%
  mutate(
    Award_Amount = as.numeric(gsub("[^0-9.-]", "", Award_Amount)),
    Start_Date = as.Date(Start_Date, format = "%m/%d/%Y"),
    End_Date = as.Date(End_Date, format = "%m/%d/%Y"),
    Funding_Agency = "NSF"
  )

# Prepare projects_df with corrected date handling and Funding_Agency
projects_df_prepared <- projects_df %>%
  select(
    PI_Name = contact_pi_name, 
    Award_Amount = award_amount, 
    Start_Date = project_start_date, 
    End_Date = project_end_date
  ) %>%
  mutate(
    Award_Amount = as.numeric(gsub("[^0-9.-]", "", Award_Amount)),
    Start_Date = as.Date(Start_Date, format = "%m/%d/%Y"),
    End_Date = as.Date(End_Date, format = "%m/%d/%Y"),
    Funding_Agency = "NIH"
  )

# Combine the prepared datasets into UI_Grant_Data
UI_Grant_Data <- bind_rows(DOE_UI_prepared, NSFtoUI_prepared, projects_df_prepared)
```

```{r}
head(UI_Grant_Data)
```





