---
title: "Mid-Term Draft 03"
author: "Konrad Bailey"
date: "2024-03-21"
output: pdf_document
eval: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Loading libraries
```{r}
library(readxl)
library(httr)
library(jsonlite)
library(tidyverse)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(ggalt)
library(tibble)
library(gridExtra)
library(stringr)
library(grid)
library(scales)
```


## Department of Energy Data

```{r}
DOEawards.data <- read.csv("~/Desktop/DOEawards.csv")
```

#### Filtering city by Moscow in DOE data
```{r}
DOE_UI.data <- DOEawards.data %>%
  filter(City == 'Moscow')
```

```{r}
head(DOE_UI.data)
```


## National Science Foundation Data

```{r}
library(readxl)


# Base URL for the API
base_url <- "https://www.research.gov/awardapi-service/v1/awards.json?awardeeName=%22regents+of+the+university+of+idaho%22"

printFields <- "rpp,offset,id,agency,awardeeCity,awardeeCountryCode,awardeeDistrictCode,awardeeName,awardeeStateCode,awardeeZipCode,cfdaNumber,coPDPI,date,startDate,expDate,estimatedTotalAmt,fundsObligatedAmt,ueiNumber,fundProgramName,parentUeiNumber,pdPIName,perfCity,perfCountryCode,perfDistrictCode,perfLocation,perfStateCode,perfZipCode,poName,primaryProgram,transType,title,awardee,poPhone,poEmail,awardeeAddress,perfAddress,publicationResearch,publicationConference,fundAgencyCode,awardAgencyCode,projectOutComesReport,abstractText,piFirstName,piMiddeInitial,piLastName,piEmail"

# Initialize an empty data frame to store results
NSFtoUI <- tibble()

# Number of results per page (as per API settings)
results_per_page <- 25

# Variable to keep track of the current page number
current_page <- 1

# Variable to control the loop
keep_going <- TRUE

while(keep_going) {
    # Calculate the offset for the current page
    offset <- (current_page - 1) * results_per_page + 1

    # Construct the full URL with offset
    url <- paste0(base_url, "&offset=", offset, "&printFields=", printFields)

    # Make the API call
    response <- GET(url)

    # Check if the call was successful
    if (status_code(response) == 200) {
        # Extract and parse the JSON data
        json_data <- content(response, type = "text", encoding = "UTF-8")
        parsed_data <- fromJSON(json_data, flatten = TRUE)

        # Extract the 'award' data and add to the all_awards data frame
        awards_data <- parsed_data$response$award
        NSFtoUI <- bind_rows(NSFtoUI, as_tibble(awards_data))

        # Debug: Print the current page number and number of awards fetched
        print(paste("Page:", current_page, "- Awards fetched:", length(awards_data$id)))

        # Check if the current page has less than results_per_page awards, then it's the last page
        if (length(awards_data$id) < results_per_page) {
            keep_going <- FALSE
        } else {
            current_page <- current_page + 1
        }
    } else {
        print(paste("Failed to fetch data: Status code", status_code(response)))
        keep_going <- FALSE
    }
}
```

```{r}
head(NSFtoUI)
```


## National Institute of Health Data

```{r}
library(readxl)


# Base URL for the API
base_url <- "https://www.research.gov/awardapi-service/v1/awards.json?awardeeName=%22regents+of+the+university+of+idaho%22"

printFields <- "rpp,offset,id,agency,awardeeCity,awardeeCountryCode,awardeeDistrictCode,awardeeName,awardeeStateCode,awardeeZipCode,cfdaNumber,coPDPI,date,startDate,expDate,estimatedTotalAmt,fundsObligatedAmt,ueiNumber,fundProgramName,parentUeiNumber,pdPIName,perfCity,perfCountryCode,perfDistrictCode,perfLocation,perfStateCode,perfZipCode,poName,primaryProgram,transType,title,awardee,poPhone,poEmail,awardeeAddress,perfAddress,publicationResearch,publicationConference,fundAgencyCode,awardAgencyCode,projectOutComesReport,abstractText,piFirstName,piMiddeInitial,piLastName,piEmail"

# Initialize an empty data frame to store results
NSFtoUI <- tibble()

# Number of results per page (as per API settings)
results_per_page <- 25

# Variable to keep track of the current page number
current_page <- 1

# Variable to control the loop
keep_going <- TRUE

while(keep_going) {
    # Calculate the offset for the current page
    offset <- (current_page - 1) * results_per_page + 1

    # Construct the full URL with offset
    url <- paste0(base_url, "&offset=", offset, "&printFields=", printFields)

    # Make the API call
    response <- GET(url)

    # Check if the call was successful
    if (status_code(response) == 200) {
        # Extract and parse the JSON data
        json_data <- content(response, type = "text", encoding = "UTF-8")
        parsed_data <- fromJSON(json_data, flatten = TRUE)

        # Extract the 'award' data and add to the all_awards data frame
        awards_data <- parsed_data$response$award
        NSFtoUI <- bind_rows(NSFtoUI, as_tibble(awards_data))

        # Debug: Print the current page number and number of awards fetched
        print(paste("Page:", current_page, "- Awards fetched:", length(awards_data$id)))

        # Check if the current page has less than results_per_page awards, then it's the last page
        if (length(awards_data$id) < results_per_page) {
            keep_going <- FALSE
        } else {
            current_page <- current_page + 1
        }
    } else {
        print(paste("Failed to fetch data: Status code", status_code(response)))
        keep_going <- FALSE
    }
}
```

```{r}
print(projects_df)
```

## Generating UI_Grant_Data

```{r}
# Prepare DOE_UI.data with Funding_Agency
DOE_UI_prepared <- DOE_UI.data %>%
  select(
    PI_Name = PI, 
    Award_Amount = Amount.Awarded.to.Date, 
    Start_Date = Start.Date, 
    End_Date = End.Date
  ) %>%
  mutate(
    Award_Amount = as.numeric(gsub("[^0-9.-]", "", Award_Amount)),
    Start_Date = mdy(Start_Date),
    End_Date = mdy(End_Date),
    Funding_Agency = "DOE"
  )

# Prepare NSFtoUI with Funding_Agency
NSFtoUI_prepared <- NSFtoUI %>%
  select(
    PI_Name = pdPIName, 
    Award_Amount = estimatedTotalAmt, 
    Start_Date = startDate, 
    End_Date = expDate
  ) %>%
  mutate(
    Award_Amount = as.numeric(gsub("[^0-9.-]", "", Award_Amount)),
    Start_Date = mdy(Start_Date),
    End_Date = mdy(End_Date),
    Funding_Agency = "NSF"
  )

# Prepare projects_df with corrected date handling for ISO 8601 format and Funding_Agency
projects_df_prepared <- projects_df %>%
  select(
    PI_Name = contact_pi_name, 
    Award_Amount = award_amount, 
    Start_Date = project_start_date, 
    End_Date = project_end_date
  ) %>%
  mutate(
    Award_Amount = as.numeric(gsub("[^0-9.-]", "", Award_Amount)),
    Start_Date = as.Date(ymd_hms(Start_Date)),
    End_Date = as.Date(ymd_hms(End_Date)),
    Funding_Agency = "NIH"
  )

# Combine the prepared datasets into UI_Grant_Data
UI_Grant_Data <- bind_rows(DOE_UI_prepared, NSFtoUI_prepared, projects_df_prepared)
```


```{r}
print(UI_Grant_Data)
```



```{r}

# Prepare DOE_UI.data with Funding_Agency
DOE_UI_prepared <- DOE_UI.data %>%
  select(
    PI_Name = PI, 
    Award_Amount = Amount.Awarded.to.Date, 
    Start_Date = Start.Date, 
    End_Date = End.Date
  ) %>%
  mutate(
    Award_Amount = as.numeric(gsub("[^0-9.-]", "", Award_Amount)),
    Start_Date = mdy(Start_Date),
    End_Date = mdy(End_Date),
    Funding_Agency = "DOE"
  )

# Prepare NSFtoUI with Funding_Agency
NSFtoUI_prepared <- NSFtoUI %>%
  select(
    PI_Name = pdPIName, 
    Award_Amount = estimatedTotalAmt, 
    Start_Date = startDate, 
    End_Date = expDate
  ) %>%
  mutate(
    Award_Amount = as.numeric(gsub("[^0-9.-]", "", Award_Amount)),
    Start_Date = mdy(Start_Date),
    End_Date = mdy(End_Date),
    Funding_Agency = "NSF"
  )

# Prepare projects_df with corrected date handling for ISO 8601 format and Funding_Agency
projects_df_prepared <- projects_df %>%
  select(
    PI_Name = contact_pi_name, 
    Award_Amount = award_amount, 
    Start_Date = project_start_date, 
    End_Date = project_end_date
  ) %>%
  mutate(
    Award_Amount = as.numeric(gsub("[^0-9.-]", "", Award_Amount)),
    Start_Date = as.Date(ymd_hms(Start_Date)),
    End_Date = as.Date(ymd_hms(End_Date)),
    Funding_Agency = "NIH"
  )

# Combine the prepared datasets into UI_Grant_Data
UI_Grant_Data <- bind_rows(DOE_UI_prepared, NSFtoUI_prepared, projects_df_prepared) %>%
  # Filter out rows where Start_Date is before 01/01/2014
  filter(Start_Date >= as.Date('2014-01-01'))

```

```{r}
print(UI_Grant_Data)
```


```{r}

# Filter the dataset for rows where Funding_Agency is "DOE"
DOE_data <- UI_Grant_Data %>% 
  filter(Funding_Agency == "DOE")

# Generate the dumbbell plot with updated font sizes
ggplot(DOE_data, aes(y = reorder(PI_Name, Start_Date), x = Start_Date, xend = End_Date)) +
  geom_dumbbell(size=1, colour="#82cbb2", colour_x="#191970", colour_xend="#191970", dot_size=4) +
  labs(title = "Timeline of Department of Energy Grants", # Title
       x = "Date", # X-axis label
       y = "Principal Investigator") + # Y-axis label
  theme_minimal() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle=45, hjust=1, size=8), # X-axis text size
        axis.text.y = element_text(size=8), # Y-axis text size
        axis.title.x = element_text(size=9.8), # X-axis label size
        axis.title.y = element_text(size=9.8), # Y-axis label size
        plot.title = element_text(face = "bold", size=12.5)) # Title font size and style

```



```{r}

# Create the dumbbell plot for the first half with adjusted axis title sizes
plot1 <- ggplot(NSF_data_part1, aes(y = reorder(PI_Name, Start_Date), x = Start_Date, xend = End_Date)) +
  geom_dumbbell(size=0.5, colour="#82cbb2", colour_x="#191970", colour_xend="#191970") +
  labs(x = "Date", y = "Principal Investigator") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle=45, hjust=1, size=7),
        axis.text.y = element_text(size=4.4),
        axis.title.x = element_text(size=8.5), # Make x-axis title smaller and match y-axis title size
        axis.title.y = element_text(size=8.5), # Adjust y-axis title size to be consistent
        panel.grid = element_blank())

# Create the dumbbell plot for the second half with adjusted axis title sizes
plot2 <- ggplot(NSF_data_part2, aes(y = reorder(PI_Name, Start_Date), x = Start_Date, xend = End_Date)) +
  geom_dumbbell(size=0.5, colour="#82cbb2", colour_x="#191970", colour_xend="#191970") +
  labs(x = "Date", y = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle=45, hjust=1, size=7),
        axis.text.y = element_text(size=4.4),
        axis.title.x = element_text(size=8.5), # Ensure x-axis title size is consistent across both plots
        panel.grid = element_blank())

# Arrange the plots side by side with a single title for the entire figure
grid.arrange(plot1, plot2, ncol = 2, top = grid::textGrob("Timeline of National Science Foundation Grants", gp = grid::gpar(fontface = "bold", fontsize = 12.5)))

```



```{r}

# Filter the dataset for rows where Funding_Agency is "NIH"
NIH_data <- UI_Grant_Data %>% 
  filter(Funding_Agency == "NIH")

# Calculate the number of rows and find the midpoint
n_rows <- nrow(NIH_data)
midpoint <- floor(n_rows / 2)

# Correctly split NIH data into two parts for plotting
NIH_data_part1 <- NIH_data %>% slice(1:midpoint)
NIH_data_part2 <- NIH_data %>% slice((midpoint + 1):n_rows)

# Create the dumbbell plot for the first half with NIH data
plot1 <- ggplot(NIH_data_part1, aes(y = reorder(PI_Name, Start_Date), x = Start_Date, xend = End_Date)) +
  geom_dumbbell(size=0.5, colour="#82cbb2", colour_x="#191970", colour_xend="#191970") +
  labs(x = "Date", y = "Principal Investigator") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle=45, hjust=1, size=7),
        axis.text.y = element_text(size=4.4),
        axis.title.x = element_text(size=8.5),
        axis.title.y = element_text(size=8.5),
        panel.grid = element_blank())

# Create the dumbbell plot for the second half with NIH data
plot2 <- ggplot(NIH_data_part2, aes(y = reorder(PI_Name, Start_Date), x = Start_Date, xend = End_Date)) +
  geom_dumbbell(size=0.5, colour="#82cbb2", colour_x="#191970", colour_xend="#191970") +
  labs(x = "Date", y = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle=45, hjust=1, size=7),
        axis.text.y = element_text(size=4.4),
        axis.title.x = element_text(size=8.5),
        panel.grid = element_blank())

# Arrange the plots side by side with a single title for the entire figure
grid.arrange(plot1, plot2, ncol = 2, top = grid::textGrob("Timeline of National Institute of Health Grants", gp = grid::gpar(fontface = "bold", fontsize = 12.5)))

```


```{r}
# Summarize the total Award_Amount for each Funding_Agency
agency_totals <- UI_Grant_Data %>%
  group_by(Funding_Agency) %>%
  summarise(Total_Award_Amount = sum(Award_Amount, na.rm = TRUE)) %>%
  arrange(desc(Total_Award_Amount)) # Optional: arrange by total amount for a more informative plot

# Create a bar chart with dollar-formatted y-axis labels and custom colors
ggplot(agency_totals, aes(x = Funding_Agency, y = Total_Award_Amount, fill = Funding_Agency)) +
  geom_bar(stat = "identity", show.legend = FALSE) + # Use stat="identity" to use the values in the data frame
  theme_minimal() +
  labs(x = "Funding Agency", y = "Total Award Amount ($)", title = "Total Award Amount by Funding Agency") +
  scale_y_continuous(labels = label_dollar()) + # Format y-axis labels as dollars
  scale_fill_manual(values = c("#7982A9", "#B0C4DE", "#507072")) + # Custom colors for each funding agency
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for better readability


```

```{r}
# Calculate the overall total Award_Amount
overall_total <- sum(UI_Grant_Data$Award_Amount, na.rm = TRUE)

# Summarize the percentage of Award_Amount for each Funding_Agency
agency_percentages <- UI_Grant_Data %>%
  group_by(Funding_Agency) %>%
  summarise(Total_Award_Amount = sum(Award_Amount, na.rm = TRUE),
            Percentage = Total_Award_Amount / overall_total * 100) %>%
  arrange(desc(Percentage)) # Optional: arrange by percentage for a more informative plot

# Create a pie chart to display the percentage contribution of each Funding_Agency
ggplot(agency_percentages, aes(x = "", y = Percentage, fill = Funding_Agency)) +
  geom_bar(width = 1, stat = "identity", show.legend = TRUE) + # Use geom_bar and set width to create a pie chart
  coord_polar(theta = "y") + # Use polar coordinates to transform the bar chart into a pie chart
  theme_void() + # Use theme_void to minimize chart junk
  labs(title = "Percentage of Total Award Amount by Funding Agency", y = "Percentage") +
  scale_fill_manual(values = c("#7982A9", "#B0C4DE", "#507072")) # Custom colors for each funding agency

```


```{r}
'#82cbb2'  "#B0C4DE" '#191970' '#527D92' '#507072' '#6B9C9D' '#5F7A8B' "#7982A9"


"#7982A9" '#82cbb2' '#5F7A8B' '#4F5B66' '#575D90'
```


```{r}
# Filter the dataset for rows where Funding_Agency is "NIH"
NIH_data <- UI_Grant_Data %>% 
  filter(Funding_Agency == "NIH")

# Create a scatter plot with Award_Amount on the x-axis and PI_Name on the y-axis
scatter_plot <- ggplot(NIH_data, aes(x = Award_Amount, y = reorder(PI_Name, Award_Amount))) +
  geom_point() +
  labs(x = "Award Amount ($)", y = "Principal Investigator") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.ticks.x = element_line(),
    axis.text.y = element_text(size = 7),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(t = 30, r = 30, b = 50, l = 50)
  ) +
  scale_x_continuous(position = "top") # Move the x-axis to the top

# Print the plot
print(scatter_plot)

```

```{r}

# Calculate the number of rows and find segment points
n_rows <- nrow(NIH_data)
thirdpoint <- floor(n_rows / 3)
two_thirdpoint <- floor(2 * n_rows / 3)

# Correctly split NIH data into three parts for plotting
NIH_data_part1 <- NIH_data %>% slice(1:thirdpoint)
NIH_data_part2 <- NIH_data %>% slice((thirdpoint + 1):two_thirdpoint)
NIH_data_part3 <- NIH_data %>% slice((two_thirdpoint + 1):n_rows)

# Create scatter plot for the first part of the data
scatter_plot1 <- ggplot(NIH_data_part1, aes(x = Award_Amount, y = reorder(PI_Name, Award_Amount))) +
  geom_point() +
  labs(x = "Award Amount ($)", y = "Principal Investigator") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    axis.text.y = element_text(size = 7),
    panel.grid.major.y = element_blank()
  )

# Create scatter plot for the second part of the data
scatter_plot2 <- ggplot(NIH_data_part2, aes(x = Award_Amount, y = reorder(PI_Name, Award_Amount))) +
  geom_point() +
  labs(x = "Award Amount ($)", y = "") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    axis.text.y = element_blank(),
    panel.grid.major.y = element_blank()
  )

# Create scatter plot for the third part of the data
scatter_plot3 <- ggplot(NIH_data_part3, aes(x = Award_Amount, y = reorder(PI_Name, Award_Amount))) +
  geom_point() +
  labs(x = "Award Amount ($)", y = "") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    axis.text.y = element_blank(),
    panel.grid.major.y = element_blank()
  )

# Arrange the three plots side by side with a single title for the entire figure
grid.arrange(scatter_plot1, scatter_plot2, scatter_plot3, ncol = 3, top = "Timeline of National Institute of Health Grants")

```

## Note:

Some PI's received multiple grants, so in certain instances there's overlapping data on single rows. I need a solution to this problem (perhaps multiple rows for PI's received multiple grants).

```{r}



```


```{r}
head(UI_Grant_Data)
```

